---
title: "Homework Assignment 9"
author: "Armen Mkrtumyan"
date: "16/04/23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Write your solutions and comments in this markdown file and submit it into moodle.

## Part 1: Basic Operations with R (50 points)

**1.1.** Create a vector *numbers* that contains numbers from 1 to 20. (5 points)

```{r}

numbers <- c(1:20)
print(numbers)
```

**1.2.** Create a vector of natural numbers from 21 to 27 (both ends inclusive) and letters "A", "B", "C". After check the data type of the vector and explain. (5 points)

```{r}
new_vector <- c(21:27, "A", "B", "C")
print(new_vector)
print(class(new_vector))
#We can see that the output is `character` 
#since when different data types are given to vector,
#R coerces the types into one common.


```

**1.3.** Create a vector "numbers_2", containing integers from 1 to 100 and 200 to 300. Print the length. (5 points)

```{r}

numbers_2 <- c(1:100, 200:300)
print(length(numbers_2))

```

**1.4.** Create another vector *odd_numbers* that contains only odd numbers from 1 to 20 using function *seq()*. (5 points)\
Hint: You can see the documentation of the function by typing *?seq*, pay attention to the argument *by*.

```{r}

odd_numbers = seq(1, 20, 2)
print(odd_numbers)
```

**1.5.** Subset the vector *odd_numbers* so that only numbers greater than 10 remain in the new vector. Name the new vector as you wish. The resulting vector should have 5 elements, check the length. (5 points)

```{r}
sub_odd = odd_numbers[odd_numbers > 10]
print(length(sub_odd))

```

**1.6.** Consider the vector given below. Merge this vector with the vector obtained in the previous point. Calculate the sum, mean, standard deviation for the obtained vector. (5 points)

```{r}
v1 <- c(12, 3, 11, 8, 4, 7, 5)

merged_vector = c(v1, sub_odd)
paste(c("sum is: ", sum(merged_vector)), collapse = " ")
paste(c("mean is: ", mean(merged_vector)), collapse = " ")
paste(c("mean is: ", sd(merged_vector)), collapse = " ")
```

**1.7.** Build a 4x3 matrix and populate it with the vector generated above, where matrix elements are filled by row. (5 points)

```{r}

my_matrix = matrix(data = merged_vector, nrow = 4, ncol = 3, byrow = T)
print(my_matrix)
```

**1.8.** Choose element from forth row and second column assigning it to a new variable *num_1*. (5 points)

```{r}

num_1 = my_matrix[4, 2]
print(num_1)

```

**1.9.** Make the rownames of matrix the followings respectively: first_row, second_row, third_row, forth_row. (5 points)

```{r}

rownames(my_matrix) <- c("first_row", "second_row", "third_row", "forth_row")
print(my_matrix)
```

**1.10.** For each row calculate the sum of elements in that row. (5 points)\
Hint: During the lecture you used *colSums* function, look at *?rowSums*.

```{r}

print(my_matrix)

print(rowSums(my_matrix))
```

## Part 2: Working with data (50 points)

**2.1.** Run the previous lines to have the dataframe. State the number of observations and features. Use *str* function. (5 points)

```{r}
df <- data.frame("YEARS_ACTIVE" = c(2001:2010), 
                 "SCORED_GOALS" = c(20, 14, 31, 11, 21, 22, 19, 19, 15, 11), 
                 "PLAYED_GAMES" = c(6, 7, 10, 3, 4, 7, 7, 7, 6, 6))
print(str(df))
```

**2.2.** The column *SCORED_GOALS* shows the goals. Find the maximum, minimum and mean values for scored goals. (10 points)

```{r}
paste(c("max is: ", max(df["SCORED_GOALS"])), collapse = " ")
paste(c("min is: ", min(df["SCORED_GOALS"])), collapse = " ")
paste(c("mean is: ", mean(df$SCORED_GOALS)), collapse = " ")
```

**2.3.** Subset the dataframe to have the columns *YEARS_ACTIVE* and *PLAYED_GAMES*. Store it in variable *df_subset*. (8 points)

```{r}

df_subset = df[c("YEARS_ACTIVE", "PLAYED_GAMES")]
print(df_subset)
```

**2.4.** Subset the dataframe to have only the rows when YEARS_ACTIVE equals 2004, 2007 and 2010, and exclude columns SCORED_GOALS and PLAYED_GAMES. (12 points)

```{r}
df[df$YEARS_ACTIVE %in% c(2002, 2004, 2007, 2010), !(names(df) %in% c("SCORED_GOALS", "PLAYED_GAMES"))]
```

**2.5.** Write a for statement that will compute the sum of even numbers from 1 to 11. Note that modulo operater in R that checks remainder after a division is "%%" rather than "%". (15 points)

```{r}

sum = 0
for (num in 1:11)
{
  if(num %% 2 == 0){
    sum = sum + num
  }
}
print(sum)

```
