---
title: "Homework Assignment 10"
author: "Armen Mkrtumyan"
date: "4/17/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1 (35 points / 5 each)

**1.1.** Read *movies3.csv* to R. See the structure of the dataframe, what are the observations and what are the features of the dataframe. Print the head of the dataset. 
```{r}
data = read.csv("movies3.csv")
print(str(data)) #We can see 2912 observations and 33 features(title, genre_first...)
```
```{r}
print(head(data))
```
**1.2.** Calculate the average of the column budget. 
```{r}
mean_budget = mean(data$budget)
print(mean_budget)
```


**1.3.** Subset the dataframe to have only columns gross_adjusted, budget_adjusted, gross, budget (store in variable movies_sub). 

```{r}
movies_sub = data[,c("gross_adjusted", "budget_adjusted", "gross", "budget")]
```

**1.4.** Calculate the mean of each column of the subsetted dataframe by *apply*.
```{r}
apply(movies_sub, MARGIN = 2, mean)
```

**1.5.** Find the movie title with the minimum badget according to the dataset.
```{r}
index = which.min(data$budget)
data[index, "title"]
```

**1.6.** Calculate how many movies are there for which the budget of the movie is smaller than the mean budget and when number of reviews is greater than 200.


```{r}
sum(data$budget < mean_budget & data$reviews > 200)
```


**1.7.** Add a new column to dataframe *Years_after_prod* that will be equal to *Release_year* - *year*.  Look at the summary of the new column.
```{r}
data$Years_after_prod <- data$Release_year - data$year
summary(data$Years_after_prod)
```


## Exercise 2 (35 points / 5 each)


**2.1.** For the following part, import *countries.csv* dataset to R by reading it. Then print first 4 observations in the dataset. Next state how many observations and how many features the dataframe has? Are all features numeric ones? 

```{r}

countries = read.csv("countries.csv")
print(head(countries, n = 4))

```


```{r}
print(str(countries)) #NO, all the features are not numeric, we also have chr, int
```


**2.2.** The column Population shows the population for all the countries. Find the maximum and minimum population values in the dataset. 


```{r}
minimum_pop = min(countries$Population)
maximum_pop = max(countries$Population)
paste("Minimum: ", minimum_pop)
paste("Maximum: ", maximum_pop)
```

**2.3.** Now when you have the minimum and maximum values of the population you can subset the dataset, so you can see which are the countries with these populations. Find those countries. 

```{r}
print(subset(countries, Population == minimum_pop))
print(subset(countries, Population == maximum_pop))

```

**2.4.** Now suppose we want to consider only those countries, which are in the region *C.W. OF IND. STATES*. Subset the dataframe as follows. Name the new dataframe *CIS_countries*. These are countries that are members of Commonwealth of Independent States (CIS) and one of them is Armenia. How many countries are there that belong to the CIS.


```{r}
CIS_countries <- subset(countries, countries$Region == "C.W. OF IND. STATES")
print(nrow(CIS_countries)) #There are 12 countries which belong to CIS
```


**2.5.** Consider the climate of the CIS countries. Calculate the mean and standard deviation of the feature. What can you say about the results? 
Hint: If you are getting NA after running functions, one reason can be that the variable has NA inside. Look at the help of the functions by typing *?function_name*, specifically for the argument *na.rm*.

```{r}
#The standard deviation of 2.55 and mean of 1.25 indicate that there is large amount of variability in the data. This means that the data points are far away from the mean and are spread out rather than being clustered. 
sd(CIS_countries$Climate, na.rm = T)
mean(CIS_countries$Climate, na.rm = T)
```
  

**2.6.** The difference between the birth rate and the death rate of a country or place is called the natural increase. The natural increase is calculated by subtracting the death rate from the birth rate. *Natural increase = birth rate - death rate.* Calculate Natural Increase and keep it in the dataframe under a column *NaturalIncrease*. Then find which countries have the highest and lowest natural increase. Are they the same as those countries with minimum and maximum Population that you found in one of previous exercises.  (6 points)  
Hint: For adding new feature, recall how you calculated number of lost games during the lectures, when you had only wins and draws. The steps for min and max should be the same as for finding the countries as they were in the case of Population. If you obtain NAs, consider what have you done while calculating mean and standard deviation to avoid the problem.

```{r}
countries$NaturalIncrease <- countries$Birthrate - countries$Deathrate
minimum =  min(countries$NaturalIncrease, na.rm = T)
maximum = max(countries$NaturalIncrease, na.rm = T)
paste("Minimum natural increase: ", subset(countries, NaturalIncrease == minimum)$Country)
paste("Maximum natural increase: ", subset(countries, NaturalIncrease == maximum)$Country)
```


**2.7.** Plot a scatterplot using ggplot2 library between *Population* and *Area_sqm*. Do not forget give proper names to axes and to the graph itself. Interpret graph in a few words. Can you make the graph any better? 
Hint: ?xlim and ?ylim
```{r}
library("ggplot2")

ggplot(data=countries, aes(x=Population, y=Area_sqm)) +geom_point() + labs(y= "Area per square meter", x = "Population") + ggtitle("Population vs Area") + coord_cartesian(xlim = c(0, 30000000), ylim = c(0, 1000000))
#The coord_caresian is added to keep all the points and not get a warning that some points were deleted due to zooming
#The graph is scattered so it does not show any linear correlation between the population per area, only in the beginning we can see
#Some relations, when fewer the population meant smaller area
```


## Exercise 3 (30 points / 10 each)

**3.1.** Using a for loop get the factorial of 14, and check your result with R's built-in function *factorial()*.

```{r}

my_factorial<-function(n)
{
  if(n == 0)
    return(1)
  else
    return(n * factorial(n - 1))
}
my_factorial(12)
factorial(12)
```



**3.2.** Write a while loop that prints out standard random normal numbers (use rnorm()) but stops (breaks) if you get a number bigger than 1.
```{r}

number <- 0
while(number <= 1)
{
  number <- rnorm(1)
  print(number)
}
print(number)

```



**3.3.** Write a function that will get an input n (n > 1), and will return the n-th Fibonacci number.
```{r}
my_fibonacci<-function(n)
{
  if(n < 2)
    return(n)
  else
    return(my_fibonacci(n-1) + my_fibonacci(n-2))
}
my_fibonacci(10)
```







